Live Functional Programming with Typed Holes

### 2018-10-23 calculus.tex
close #17

The progress theorem establishes that the dynamic semantics accounts
for every well-typed term, i.e. that we have not forgotten some
necessary rules or premises.
%
\begin{thm}[Progress]
  If $\hasType{\Delta}{\emptyset}{\dexp}{\htau}$ then either
  (a) $\stepsToD{}{\dexp}{\dexp'}$ or
  (b) $\isBoxedValue{\dexp}$ or
  (c) $\isIndet{\dexp}$.
\end{thm}
\noindent
The key to establishing the progress theorem under a non-empty hole
context is to explicitly account for indeterminate forms,
i.e. those rooted at either a hole closure or a failed cast.
%
The proof relies on canonical forms lemmas stated in the \appendixName.
% Appendix~\ref{sec:additional-defns}.

@@

The following progress theorem establishes that evaluating a complete
program always results in classic values, not boxed values nor
indeterminate forms.
%
\begin{thm}[Complete Progress]
  If $\hasType{\hDelta}{\emptyset}{\dexp}{\htau}$
  and $\isComplete{\dexp}$
  then either $\stepsToD{}{\dexp}{\dexp'}$
  or $\isValue{\dexp}$.
\end{thm}

>>>>

The progress theorem establishes that the dynamic semantics accounts
for every well-typed term, i.e. that we have not forgotten some
necessary rules or premises.
%
\begin{thm}[Progress]
  If $\hasType{\Delta}{\emptyset}{\dexp}{\htau}$ then either
  (a) there exists $\dexp'$ such that $\stepsToD{}{\dexp}{\dexp'}$ or
  (b) $\isBoxedValue{\dexp}$ or
  (c) $\isIndet{\dexp}$.
\end{thm}
\noindent
The key to establishing the progress theorem under a non-empty hole
context is to explicitly account for indeterminate forms,
i.e. those rooted at either a hole closure or a failed cast.
%
The proof relies on canonical forms lemmas stated in the \appendixName.
% Appendix~\ref{sec:additional-defns}.

@@

The following progress theorem establishes that evaluating a complete
program always results in classic values, not boxed values nor
indeterminate forms.
%
\begin{thm}[Complete Progress]
  If $\hasType{\hDelta}{\emptyset}{\dexp}{\htau}$ and $\isComplete{\dexp}$
  then either there exists a $\dexp'$ such that
  $\stepsToD{}{\dexp}{\dexp'}$, or $\isValue{\dexp}$.
\end{thm}
